\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.7in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftTint}{HTML}{F7F4EE}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftGreen}{HTML}{EEF7F0}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Paragraph Workbook - 2026-04-27},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.25}
\setlist[itemize]{leftmargin=1.3em,itemsep=2pt,topsep=3pt}
\setlist[enumerate]{leftmargin=1.45em,itemsep=3pt,topsep=4pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\Large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.1em}{\color{AccentGold}\titlerule[0.8pt]}]

\titleformat{\subsection}
  {\normalfont\large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}April 27 Workbook}
\fancyhead[R]{\sffamily\color{HeadingBlue}IB Geography SL}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{AccentTeal}{\textbf{#1}}}
\newcommand{\todobox}{\fbox{\rule{0pt}{0.8em}\rule{0.8em}{0pt}}}
\newcommand{\blankbox}[2]{%
\noindent\fcolorbox{RuleGray}{#1}{%
  \begin{minipage}[t][#2][t]{0.965\textwidth}
  \vspace{0.2em}
  \end{minipage}
}}

\begin{document}

\begin{center}
{\sffamily\bfseries\color{HeadingBlue}\LARGE April 27 Paragraph Practice Workbook}\\[0.25em]
{\sffamily\color{AccentTeal}\large Urban Processes + Climate Vulnerability And Resilience}
\end{center}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Use with the April 27 study pack.}}
The plan requires one \exam{explain} paragraph and one \exam{analyse}
paragraph in \term{25 minutes}. Work from memory first, then check the study
pack and correct in a different color.
\end{minipage}
}

\section{Session Checklist}

\begin{tabularx}{\textwidth}{L{0.12\textwidth}Y L{0.18\textwidth}}
\toprule
\textbf{Done} & \textbf{Task} & \textbf{Time} \\
\midrule
\todobox & Closed-book brain dump: urban process chain and climate risk chain & 5 min \\
\todobox & Write one \exam{explain} paragraph & 8 min \\
\todobox & Write one \exam{analyse} paragraph & 8 min \\
\todobox & Self-mark, fix, and log next action & 4 min \\
\bottomrule
\end{tabularx}

\section{Minute 0--5: Closed-Book Brain Dump}

\subsection{Urban Process Chain}

Fill the chain from memory before reading the prompts.

\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y}
\toprule
\textbf{Step} & \textbf{My notes} \\
\midrule
Cause of urbanization & \rule{0.78\linewidth}{0.4pt} \\
Movement or growth process & \rule{0.78\linewidth}{0.4pt} \\
Urban pattern created & \rule{0.78\linewidth}{0.4pt} \\
Infrastructure demand & \rule{0.78\linewidth}{0.4pt} \\
Impact on people or environment & \rule{0.78\linewidth}{0.4pt} \\
Evaluation or limitation & \rule{0.78\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\subsection{Climate Risk Chain}

\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y}
\toprule
\textbf{Step} & \textbf{My notes} \\
\midrule
Human cause & \rule{0.78\linewidth}{0.4pt} \\
Physical climate change & \rule{0.78\linewidth}{0.4pt} \\
Specific hazard or stress & \rule{0.78\linewidth}{0.4pt} \\
Vulnerable group or place & \rule{0.78\linewidth}{0.4pt} \\
Impact & \rule{0.78\linewidth}{0.4pt} \\
Resilience or adaptation response & \rule{0.78\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{Minute 5--13: Explain Paragraph}

\noindent\colorbox{SoftTint}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt.}}
\exam{Explain} how urbanization can create pressure on urban infrastructure.
\end{minipage}
}

\subsection{Planning Box}

\begin{tabularx}{\textwidth}{L{0.2\textwidth}Y}
\toprule
\textbf{Need} & \textbf{Write a brief note before the paragraph} \\
\midrule
Definition or opening idea & \rule{0.78\linewidth}{0.4pt} \\
Cause of growth & \rule{0.78\linewidth}{0.4pt} \\
Infrastructure example & \rule{0.78\linewidth}{0.4pt} \\
Impact & \rule{0.78\linewidth}{0.4pt} \\
Named place if known & \rule{0.78\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\subsection{Paragraph Space}

\blankbox{white}{8.2cm}

\subsection{Explain Self-Check}

\begin{tabularx}{\textwidth}{L{0.1\textwidth}Y}
\toprule
\textbf{Check} & \textbf{Question} \\
\midrule
\todobox & Did I give a clear reason, not just a description? \\
\todobox & Did I show a causal chain using words such as because, therefore,
this leads to, or as a result? \\
\todobox & Did I name at least two infrastructure systems, such as transport,
housing, water, sanitation, energy, healthcare, or schools? \\
\todobox & Did I avoid saying that population growth alone explains everything? \\
\bottomrule
\end{tabularx}

\newpage

\section{Minute 13--21: Analyse Paragraph}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt.}}
\exam{Analyse} why the consequences of climate change are uneven between
different places or groups.
\end{minipage}
}

\subsection{Planning Box}

\begin{tabularx}{\textwidth}{L{0.21\textwidth}Y}
\toprule
\textbf{Need} & \textbf{Write a brief note before the paragraph} \\
\midrule
Physical hazard or stress & \rule{0.76\linewidth}{0.4pt} \\
Exposed place or group & \rule{0.76\linewidth}{0.4pt} \\
Sensitivity factor & \rule{0.76\linewidth}{0.4pt} \\
Resilience factor & \rule{0.76\linewidth}{0.4pt} \\
Comparison or contrast & \rule{0.76\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\subsection{Paragraph Space}

\blankbox{white}{8.6cm}

\subsection{Analyse Self-Check}

\begin{tabularx}{\textwidth}{L{0.1\textwidth}Y}
\toprule
\textbf{Check} & \textbf{Question} \\
\midrule
\todobox & Did I break the issue into parts: hazard, exposure, vulnerability,
resilience, and outcome? \\
\todobox & Did I compare at least two places, groups, or levels of capacity? \\
\todobox & Did I use \term{vulnerability} or \term{resilience} correctly? \\
\todobox & Did I explain why the impact varies rather than only listing
climate impacts? \\
\bottomrule
\end{tabularx}

\section{Minute 21--25: Mark And Fix}

\subsection{Quick Score}

\begin{tabularx}{\textwidth}{L{0.24\textwidth}L{0.17\textwidth}Y}
\toprule
\textbf{Paragraph} & \textbf{Score} & \textbf{Main reason} \\
\midrule
Explain paragraph & \rule{0.08\textwidth}{0.4pt} / 5 & \rule{0.9\linewidth}{0.4pt} \\
Analyse paragraph & \rule{0.08\textwidth}{0.4pt} / 5 & \rule{0.9\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\subsection{Fix One Sentence}

Choose the weakest sentence from either paragraph and rewrite it so it better
matches the command term.

\begin{tabularx}{\textwidth}{Y}
\toprule
\textbf{Original weak sentence} \\
\midrule
\rule{0pt}{1.3cm} \\
\bottomrule
\end{tabularx}

\vspace{0.65em}

\begin{tabularx}{\textwidth}{Y}
\toprule
\textbf{Improved sentence} \\
\midrule
\rule{0pt}{1.3cm} \\
\bottomrule
\end{tabularx}

\subsection{Error Log}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.2\textwidth}Y Y}
\toprule
\textbf{Weakness} & \textbf{What I need to fix} & \textbf{Next action} \\
\midrule
Command term & & \\
Causal chain & & \\
Case-study evidence & & \\
Climate vocabulary & & \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\newpage

\section{Support Sheet}

\subsection{Command-Term Reminders}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y Y}
\toprule
\textbf{Command} & \textbf{What it requires} & \textbf{Useful sentence moves} \\
\midrule
Explain & Reasons and causal links & This occurs because...; This leads to...;
As a result...; This creates pressure on... \\
Analyse & Break into parts and show relationships & The impact varies because...;
This depends on...; In contrast...; Therefore the outcome is uneven... \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsection{Useful Vocabulary Bank}

\begin{tabularx}{\textwidth}{Y Y}
\toprule
\textbf{\textcolor{HeadingBlue}{Urban processes}} &
\textbf{\textcolor{HeadingBlue}{Climate risk}} \\
\midrule
urbanization; rural-to-urban migration; natural increase; centripetal
movement; centrifugal movement; suburbanization; reurbanization;
gentrification; infrastructure; congestion; sanitation; informal settlement;
land value; planning capacity &
enhanced greenhouse effect; greenhouse gases; fossil fuels; deforestation;
sea-level rise; drought; flood risk; heat stress; food security; water
security; vulnerability; exposure; sensitivity; resilience; mitigation;
adaptation \\
\bottomrule
\end{tabularx}

\subsection{Case-Study Gap Check}

The student case-study sheet has an urban anchor, \term{gentrification in
Detroit}, but the climate section is still blank. Fill this before the later
climate sessions.

\begin{tabularx}{\textwidth}{L{0.23\textwidth}Y Y}
\toprule
\textbf{Topic} & \textbf{Minimum case-study detail} & \textbf{My class example} \\
\midrule
Urban process & Place, process, infrastructure or land-use impact, stakeholder,
evaluation & \rule{0.42\linewidth}{0.4pt} \\
Climate impact & Place, hazard, exposed group, impact, why vulnerable &
\rule{0.42\linewidth}{0.4pt} \\
Climate resilience & Place, response, who benefits, limitation &
\rule{0.42\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{Mastery Update}

Rate yourself after marking. Be strict enough that the score is useful.

\begin{tabularx}{\textwidth}{Y L{0.15\textwidth}Y}
\toprule
\textbf{Skill} & \textbf{Rating} & \textbf{Evidence} \\
\midrule
I can distinguish centripetal and centrifugal movement. & \_\_ /5 & \\
I can explain urbanization as a process chain. & \_\_ /5 & \\
I can link urban growth to infrastructure pressure. & \_\_ /5 & \\
I can explain the enhanced greenhouse effect. & \_\_ /5 & \\
I can use vulnerability and resilience correctly. & \_\_ /5 & \\
I can write one explain and one analyse paragraph under time. & \_\_ /5 & \\
\bottomrule
\end{tabularx}

\vspace{0.7em}

\noindent\colorbox{SoftGreen}{%
\begin{minipage}{0.965\textwidth}
\sffamily\raggedright
\textbf{\textcolor{HeadingBlue}{Carry forward to Tuesday 2026-04-28.}}
My weakest paragraph skill is:
\rule{0.52\textwidth}{0.4pt}
\end{minipage}
}

\end{document}
