\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.68in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage[normalem]{ulem}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftTint}{HTML}{F7F4EE}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftGreen}{HTML}{EEF7F0}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Paper 2 Simulation Answer Booklet - 2026-05-03},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.38em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.17}
\setlist[enumerate]{leftmargin=1.45em,itemsep=3pt,topsep=4pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\Large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.08em}{\color{AccentGold}\titlerule[0.8pt]}]

\titleformat{\subsection}
  {\normalfont\large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}

\titlespacing*{\section}{0pt}{0.86em}{0.3em}
\titlespacing*{\subsection}{0pt}{0.58em}{0.15em}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}May 3 Answer Booklet}
\fancyhead[R]{\sffamily\color{HeadingBlue}IB Geography SL Paper 2}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\todobox}{\fbox{\rule{0pt}{0.8em}\rule{0.8em}{0pt}}}
\newcommand{\blankbox}[2]{%
\noindent\fcolorbox{RuleGray}{#1}{%
  \begin{minipage}[t][#2][t]{0.965\textwidth}
  \vspace{0.16em}
  \end{minipage}
}}

\begin{document}

\begin{center}
{\sffamily\bfseries\color{HeadingBlue}\LARGE May 3 Paper 2 Simulation Answer Booklet}\\[0.25em]
{\sffamily\color{AccentTeal}\large Population + Climate + Resources}
\end{center}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Use under timed conditions.}}
Start the 75-minute timer before writing. Keep the marking guide closed until
the timer ends. Move on when a question's time is over.
\end{minipage}
}

\section{Before You Start}

\begin{tabularx}{\textwidth}{L{0.16\textwidth}Y L{0.19\textwidth}}
\toprule
\textbf{Done} & \textbf{Action} & \textbf{Time} \\
\midrule
\todobox & Notes, phone, and marking guide put away & Before timer \\
\todobox & Read all source headings and Section C choices & 2--3 min \\
\todobox & Write start time and finish time & 1 min \\
\bottomrule
\end{tabularx}

\vspace{0.3em}

\begin{tabularx}{\textwidth}{L{0.24\textwidth}Y Y}
\toprule
\textbf{Start time} & \textbf{Section C start target} & \textbf{Finish time} \\
\midrule
\rule{0pt}{1cm} & & \\
\bottomrule
\end{tabularx}

\newpage

\section{Section A: Core Structured Questions}

\subsection{Question 1(a)}
\textbf{State the country in Source A with the oldest population structure.}
\hfill \textbf{[1]}

\blankbox{white}{1.6cm}

\subsection{Question 1(b)}
\textbf{Using Source A, describe two contrasts between Country Alpha and
Country Beta.} \hfill \textbf{[4]}

\blankbox{white}{5.0cm}

\subsection{Question 1(c)}
\textbf{Explain two possible consequences of either a youthful or an ageing
population structure.} \hfill \textbf{[4]}

\blankbox{white}{6.4cm}

\subsection{Question 1(d)}
\textbf{Suggest one reason why Country Gamma's population may continue growing
even though its fertility rate is close to replacement level.} \hfill
\textbf{[1]}

\blankbox{white}{1.8cm}

\newpage

\subsection{Question 2(a)}
\textbf{State two climate hazards shown in Source B.} \hfill \textbf{[2]}

\blankbox{white}{2.5cm}

\subsection{Question 2(b)}
\textbf{Using Source B, describe two differences in vulnerability or resilience
between the low-lying delta coast and the high-income coastal city.} \hfill
\textbf{[4]}

\blankbox{white}{5.2cm}

\subsection{Question 2(c)}
\textbf{Explain two reasons why climate risk is not determined by physical
hazard alone.} \hfill \textbf{[4]}

\blankbox{white}{6.8cm}

\newpage

\subsection{Question 3(a)}
\textbf{Identify one indicator from Source C showing uneven resource
consumption.} \hfill \textbf{[1]}

\blankbox{white}{1.6cm}

\subsection{Question 3(b)}
\textbf{Using Source C, describe two contrasts in resource security between
the high-income consumer economy and the low-income rural economy.} \hfill
\textbf{[4]}

\blankbox{white}{5.0cm}

\subsection{Question 3(c)}
\textbf{Explain two ways resource consumption can create environmental or
geopolitical insecurity.} \hfill \textbf{[4]}

\blankbox{white}{6.5cm}

\subsection{Question 3(d)}
\textbf{Suggest one stewardship strategy that could reduce resource
insecurity.} \hfill \textbf{[1]}

\blankbox{white}{1.8cm}

\newpage

\section{Section B: Visual Stimulus Task}

\subsection{Question 4(a)}
\textbf{Using Source D, describe two patterns in population growth, climate
risk, water stress, or informal housing.} \hfill \textbf{[4]}

\blankbox{white}{5.5cm}

\subsection{Question 4(b)}
\textbf{Suggest two reasons why regions with similar climate-risk indexes may
have different human impacts.} \hfill \textbf{[4]}

\blankbox{white}{6.5cm}

\subsection{Question 4(c)}
\textbf{State one decision-maker who could use Source D and one decision they
might make.} \hfill \textbf{[2]}

\blankbox{white}{3.0cm}

\newpage

\section{Section C: Extended Answer}

\subsection{Question Choice}

Tick one only.

\begin{tabularx}{\textwidth}{L{0.1\textwidth}Y}
\toprule
\todobox & 5(a) Using examples, evaluate the success of strategies to build
resilience to climate change. \\
\todobox & 5(b) Using examples, to what extent can resource stewardship reduce
the environmental and social impacts of increasing consumption? \\
\bottomrule
\end{tabularx}

\subsection{Quick Plan}

\begin{tabularx}{\textwidth}{L{0.2\textwidth}Y}
\toprule
\textbf{Chosen question} & \uline{\hspace{0.7\linewidth}} \\
\textbf{Judgment} & \rule{0pt}{1.1cm} \\
\textbf{Examples} & \rule{0pt}{1.1cm} \\
\bottomrule
\end{tabularx}

\subsection{Extended Answer}

\textbf{Chosen question:} \uline{\hspace{3.2cm}} \hfill \textbf{[10]}

\blankbox{white}{18.4cm}

\newpage

\section{Immediate Reflection}

Fill this in before opening the marking guide. Keep it short and honest.

\begin{tabularx}{\textwidth}{L{0.3\textwidth}Y}
\toprule
\textbf{Prompt} & \textbf{My note} \\
\midrule
Hardest core unit & \rule{0.64\linewidth}{0.4pt} \\
Question that took too long & \rule{0.64\linewidth}{0.4pt} \\
Weakest command term & \rule{0.64\linewidth}{0.4pt} \\
Best named example used & \rule{0.64\linewidth}{0.4pt} \\
Case-study detail I could not remember & \rule{0.64\linewidth}{0.4pt} \\
One thing to fix on May 4 & \rule{0.64\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{Score Sheet}

Complete this only after using the marking guide.

\begin{tabularx}{\textwidth}{L{0.2\textwidth}L{0.16\textwidth}Y}
\toprule
\textbf{Question} & \textbf{Score} & \textbf{Why} \\
\midrule
Q1 Population & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.56\textwidth}{0.4pt} \\
Q2 Climate & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.56\textwidth}{0.4pt} \\
Q3 Resources & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.56\textwidth}{0.4pt} \\
Q4 Visual & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.56\textwidth}{0.4pt} \\
Q5 Extended & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.56\textwidth}{0.4pt} \\
\textbf{Total} & \rule{0.08\textwidth}{0.4pt} / 50 & \rule{0.56\textwidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{15-Minute Review Error Log}

\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y Y Y}
\toprule
\textbf{Error type} & \textbf{What I wrote or missed} & \textbf{Correct version} & \textbf{May 4 fix} \\
\midrule
Source use & \rule{0pt}{1.25cm} & & \\
Process chain & \rule{0pt}{1.25cm} & & \\
Case evidence & \rule{0pt}{1.25cm} & & \\
Evaluation & \rule{0pt}{1.25cm} & & \\
Timing & \rule{0pt}{1.25cm} & & \\
\bottomrule
\end{tabularx}

\end{document}
