\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.68in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage[normalem]{ulem}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftTint}{HTML}{F7F4EE}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftGreen}{HTML}{EEF7F0}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Paper 1 Simulation Answer Booklet - 2026-05-02},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.18}
\setlist[enumerate]{leftmargin=1.45em,itemsep=3pt,topsep=4pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\Large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.08em}{\color{AccentGold}\titlerule[0.8pt]}]

\titleformat{\subsection}
  {\normalfont\large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}

\titlespacing*{\section}{0pt}{0.9em}{0.35em}
\titlespacing*{\subsection}{0pt}{0.65em}{0.18em}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}May 2 Answer Booklet}
\fancyhead[R]{\sffamily\color{HeadingBlue}IB Geography SL}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{AccentTeal}{\textbf{#1}}}
\newcommand{\todobox}{\fbox{\rule{0pt}{0.8em}\rule{0.8em}{0pt}}}
\newcommand{\blankbox}[2]{%
\noindent\fcolorbox{RuleGray}{#1}{%
  \begin{minipage}[t][#2][t]{0.965\textwidth}
  \vspace{0.18em}
  \end{minipage}
}}

\begin{document}

\begin{center}
{\sffamily\bfseries\color{HeadingBlue}\LARGE May 2 Paper 1 Simulation Answer Booklet}\\[0.25em]
{\sffamily\color{AccentTeal}\large Urban Environments + Food And Health}
\end{center}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Use under timed conditions.}}
Start the timer before writing. Do not pause between options. If you do not
know an answer, write the best geographic chain you can and move on.
\end{minipage}
}

\section{Before You Start}

\begin{tabularx}{\textwidth}{L{0.16\textwidth}Y L{0.19\textwidth}}
\toprule
\textbf{Done} & \textbf{Action} & \textbf{Time} \\
\midrule
\todobox & Notes, phone, and mark guide put away & Before timer \\
\todobox & Read the instruction page and both option headings & 1 min \\
\todobox & Write start time and finish time & 1 min \\
\bottomrule
\end{tabularx}

\vspace{0.35em}

\begin{tabularx}{\textwidth}{L{0.24\textwidth}Y Y}
\toprule
\textbf{Start time} & \textbf{Halfway switch time} & \textbf{Finish time} \\
\midrule
\rule{0pt}{1cm} & & \\
\bottomrule
\end{tabularx}

\newpage

\section{Option G: Urban Environments}

\subsection{Question 1}

\textbf{State two pieces of evidence from Source A that show uneven
environmental quality between urban zones.} \hfill \textbf{[2]}

\blankbox{white}{3.4cm}

\subsection{Question 2}

\textbf{Describe two contrasts between the inner-city regeneration zone and the
peripheral informal settlement shown in Source A.} \hfill \textbf{[4]}

\blankbox{white}{5.5cm}

\subsection{Question 3}

\textbf{Explain two reasons why rapid urban growth can create social or
environmental stress.} \hfill \textbf{[4]}

\blankbox{white}{7.2cm}

\newpage

\subsection{Question 4 Choice}

Tick one only.

\begin{tabularx}{\textwidth}{L{0.1\textwidth}Y}
\toprule
\todobox & 4(a) Using one named city, examine the success of strategies to
manage urban social or environmental stress sustainably. \\
\todobox & 4(b) Using one or more named examples, to what extent can
regeneration or gentrification solve problems caused by urban decline? \\
\bottomrule
\end{tabularx}

\subsection{Question 4 Answer}

\textbf{Chosen question:} \uline{\hspace{3.2cm}} \hfill \textbf{[10]}

\blankbox{white}{20.8cm}

\newpage

\section{Option F: Food And Health}

\subsection{Question 1}

\textbf{State two contrasts in food and health shown in Source B.}
\hfill \textbf{[2]}

\blankbox{white}{3.4cm}

\subsection{Question 2}

\textbf{Describe two ways Source B suggests that health problems can change as
diets and food systems change.} \hfill \textbf{[4]}

\blankbox{white}{5.5cm}

\subsection{Question 3}

\textbf{Explain two reasons why stakeholders can improve or worsen food
security and health.} \hfill \textbf{[4]}

\blankbox{white}{7.2cm}

\newpage

\subsection{Question 4 Choice}

Tick one only.

\begin{tabularx}{\textwidth}{L{0.1\textwidth}Y}
\toprule
\todobox & 4(a) Using one named example, examine the role of transnational
corporations in food security and health. \\
\todobox & 4(b) Using one or more named examples, evaluate the success of
strategies to improve food security or health. \\
\bottomrule
\end{tabularx}

\subsection{Question 4 Answer}

\textbf{Chosen question:} \uline{\hspace{3.2cm}} \hfill \textbf{[10]}

\blankbox{white}{20.8cm}

\newpage

\section{Immediate Reflection}

Fill this in before opening the marking guide. Keep it short and honest.

\begin{tabularx}{\textwidth}{L{0.28\textwidth}Y}
\toprule
\textbf{Prompt} & \textbf{My note} \\
\midrule
Hardest option & \rule{0.66\linewidth}{0.4pt} \\
Question that took too long & \rule{0.66\linewidth}{0.4pt} \\
Weakest command term & \rule{0.66\linewidth}{0.4pt} \\
Best named example used & \rule{0.66\linewidth}{0.4pt} \\
Case-study detail I could not remember & \rule{0.66\linewidth}{0.4pt} \\
One thing to fix on May 4 & \rule{0.66\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{Score Sheet}

Complete this only after using the marking guide.

\begin{tabularx}{\textwidth}{L{0.18\textwidth}L{0.16\textwidth}Y}
\toprule
\textbf{Question} & \textbf{Score} & \textbf{Why} \\
\midrule
Urban Q1 & \rule{0.08\textwidth}{0.4pt} / 2 & \rule{0.58\textwidth}{0.4pt} \\
Urban Q2 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
Urban Q3 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
Urban Q4 & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.58\textwidth}{0.4pt} \\
Food Q1 & \rule{0.08\textwidth}{0.4pt} / 2 & \rule{0.58\textwidth}{0.4pt} \\
Food Q2 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
Food Q3 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
Food Q4 & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.58\textwidth}{0.4pt} \\
\textbf{Total} & \rule{0.08\textwidth}{0.4pt} / 40 & \rule{0.58\textwidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\end{document}
