\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.72in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage{tikz}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{AccentRed}{HTML}{B85C5A}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Paper 1 Full Simulation Set 2 - 2026-05-07},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.16}
\setlist[enumerate]{leftmargin=1.55em,itemsep=5pt,topsep=5pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\huge\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.12em}{\color{AccentGold}\titlerule[1pt]}]
\titleformat{\subsection}
  {\normalfont\Large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}Set 2}
\fancyhead[R]{\sffamily\color{HeadingBlue}Paper 1 Simulation}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{AccentTeal}{\textbf{#1}}}
\newcommand{\markpts}[1]{\hfill\textbf{[#1]}}
\newcommand{\boxnote}[3]{%
\noindent\colorbox{#1}{%
  \begin{minipage}{0.965\textwidth}
  \sffamily
  \textbf{\textcolor{HeadingBlue}{#2}} #3
  \end{minipage}
}}

\begin{document}

\begin{center}
  {\sffamily\bfseries\color{HeadingBlue}\LARGE IB Geography SL}\\[0.2em]
  {\sffamily\bfseries\color{AccentTeal}\Large Paper 1 Full Simulation: Set 2}\\[0.3em]
  {\sffamily\color{TextInk}Thursday 2026-05-07: Urban Environments + Food And Health}
\end{center}

\boxnote{SoftRed}{Practice-paper notice.}{This is an IB-style practice paper
created for revision. It is not an official IB past paper or official IB mark
scheme. Complete it without notes before using the marking guide.}

\section{Instructions}

\begin{tabularx}{\textwidth}{L{0.25\textwidth}Y}
\toprule
\textbf{Total time} & \term{1 hour 30 minutes}. Spend about \term{45 minutes}
on each option. \\
\textbf{Total marks} & \term{40 marks}. Urban environments is worth
\term{20 marks}; Food and health is worth \term{20 marks}. \\
\textbf{Structure} & Answer every structured question. For each option, choose
\term{one} of the two 10-mark extended-answer prompts. \\
\textbf{Materials} & Question paper, answer booklet, timer, and pen. Do not use
notes during the simulation. \\
\bottomrule
\end{tabularx}

\subsection{Timing Checkpoints}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y L{0.2\textwidth}}
\toprule
\textbf{Minute} & \textbf{Action} & \textbf{Marks done} \\
\midrule
0--2 & Read the Urban source and question choices. & 0 \\
2--20 & Answer Urban structured questions 1--3. & 10 \\
20--45 & Write one Urban 10-mark answer. & 20 \\
45--47 & Switch to Food and health. & 20 \\
47--65 & Answer Food and health structured questions 1--3. & 30 \\
65--90 & Write one Food and health 10-mark answer. & 40 \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\newpage

\section{Option G: Urban Environments}

\subsection{Source A: Urban Transport, Air Quality, And Access}

Source A shows selected indicators for four zones in a large city. The data are
illustrative and are provided for interpretation practice.

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.19\textwidth}C{0.115\textwidth}C{0.115\textwidth}C{0.115\textwidth}C{0.115\textwidth}Y}
\toprule
\textbf{Zone} & \textbf{Frequent transit access} & \textbf{Average one-way commute}
& \textbf{Air-pollution index} & \textbf{Green space per person} & \textbf{Main issue} \\
\midrule
Central high-density core & 91\% & 34 min & 72 & 5 m$^2$ & Congestion and
low green space \\
Transit corridor suburb & 74\% & 48 min & 45 & 18 m$^2$ & Good access but
rising station-area rents \\
Outer car-dependent suburb & 26\% & 67 min & 61 & 31 m$^2$ & Long commutes
and land consumption \\
Informal hillside settlement & 18\% & 59 min & 83 & 2 m$^2$ & Weak transit
and poor air quality \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsubsection{Visual Source A}

Longer bars show higher values. The pollution index runs from 0 to 100, where
higher values mean poorer air quality.

\begin{center}
\begin{tikzpicture}[x=0.085cm,y=0.53cm, every node/.style={font=\sffamily\small}]
  \node[anchor=east] at (0,6.2) {Central core};
  \draw[fill=AccentTeal!72, draw=none] (0,5.95) rectangle (91,6.35);
  \draw[fill=AccentRed!70, draw=none] (0,5.45) rectangle (72,5.85);
  \node[anchor=east] at (0,4.8) {Transit corridor};
  \draw[fill=AccentTeal!72, draw=none] (0,4.55) rectangle (74,4.95);
  \draw[fill=AccentRed!70, draw=none] (0,4.05) rectangle (45,4.45);
  \node[anchor=east] at (0,3.4) {Outer suburb};
  \draw[fill=AccentTeal!72, draw=none] (0,3.15) rectangle (26,3.55);
  \draw[fill=AccentRed!70, draw=none] (0,2.65) rectangle (61,3.05);
  \node[anchor=east] at (0,2.0) {Informal hillside};
  \draw[fill=AccentTeal!72, draw=none] (0,1.75) rectangle (18,2.15);
  \draw[fill=AccentRed!70, draw=none] (0,1.25) rectangle (83,1.65);
  \foreach \x in {0,20,40,60,80,100} {
    \draw[RuleGray] (\x,0.85) -- (\x,6.55);
    \node[anchor=north] at (\x,0.75) {\x};
  }
  \node[anchor=west, color=AccentTeal] at (104,5.9) {Transit access};
  \node[anchor=west, color=AccentRed] at (104,5.4) {Air-pollution index};
\end{tikzpicture}
\end{center}

\subsection{Questions}

Answer all parts. For Question 4, answer either \term{4(a)} or \term{4(b)}.

\begin{enumerate}
  \item \exam{State} two pieces of evidence from Source A that show uneven
  access to sustainable transport or environmental quality. \markpts{2}

  \item \exam{Describe} two contrasts between the transit corridor suburb and
  the outer car-dependent suburb. \markpts{4}

  \item \exam{Explain} two ways urban transport patterns can create social or
  environmental stress. \markpts{4}

  \item Answer either 4(a) or 4(b). \markpts{10}
  \begin{enumerate}[label=\textbf{\alph*.},leftmargin=1.6em]
    \item Using one named city, \exam{evaluate} the success of sustainable
    transport strategies in managing urban environmental stress.

    \item Using one or more named examples, \exam{to what extent} does urban
    sprawl create greater challenges than high-density urban growth?
  \end{enumerate}
\end{enumerate}

\newpage

\section{Option F: Food And Health}

\subsection{Source B: Food Security And Health After A Climate Shock}

Source B shows selected indicators for four regions after an extended dry
season and a rise in food prices. The data are illustrative.

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.20\textwidth}C{0.12\textwidth}C{0.12\textwidth}C{0.12\textwidth}C{0.12\textwidth}Y}
\toprule
\textbf{Region} & \textbf{Food-insecure households} & \textbf{Child wasting}
& \textbf{Safe water access} & \textbf{Food-price increase} & \textbf{Main health risk} \\
\midrule
Irrigated core & 12\% & 5\% & 93\% & 8\% & Lower risk, but high water use \\
Dryland smallholder zone & 46\% & 21\% & 54\% & 36\% & Undernutrition and diarrhoeal disease \\
Urban low-income district & 31\% & 9\% & 76\% & 28\% & Diet quality, food price, and sanitation stress \\
Export farming belt & 18\% & 7\% & 84\% & 17\% & Uneven access for landless workers \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsubsection{Visual Source B}

Longer bars show higher values. Safe water access is beneficial; food
insecurity and child wasting are vulnerabilities.

\begin{center}
\begin{tikzpicture}[x=0.09cm,y=0.5cm, every node/.style={font=\sffamily\small}]
  \node[anchor=east] at (0,8.1) {Irrigated core};
  \draw[fill=AccentTeal!72, draw=none] (0,7.85) rectangle (12,8.15);
  \draw[fill=AccentGold!72, draw=none] (0,7.45) rectangle (5,7.75);
  \draw[fill=HeadingBlue!45, draw=none] (0,7.05) rectangle (93,7.35);
  \node[anchor=east] at (0,6.35) {Dryland smallholder};
  \draw[fill=AccentTeal!72, draw=none] (0,6.10) rectangle (46,6.40);
  \draw[fill=AccentGold!72, draw=none] (0,5.70) rectangle (21,6.00);
  \draw[fill=HeadingBlue!45, draw=none] (0,5.30) rectangle (54,5.60);
  \node[anchor=east] at (0,4.60) {Urban low-income};
  \draw[fill=AccentTeal!72, draw=none] (0,4.35) rectangle (31,4.65);
  \draw[fill=AccentGold!72, draw=none] (0,3.95) rectangle (9,4.25);
  \draw[fill=HeadingBlue!45, draw=none] (0,3.55) rectangle (76,3.85);
  \node[anchor=east] at (0,2.85) {Export farming};
  \draw[fill=AccentTeal!72, draw=none] (0,2.60) rectangle (18,2.90);
  \draw[fill=AccentGold!72, draw=none] (0,2.20) rectangle (7,2.50);
  \draw[fill=HeadingBlue!45, draw=none] (0,1.80) rectangle (84,2.10);
  \foreach \x in {0,20,40,60,80,100} {
    \draw[RuleGray] (\x,1.35) -- (\x,8.45);
    \node[anchor=north] at (\x,1.25) {\x\%};
  }
  \node[anchor=west, color=AccentTeal] at (104,7.85) {Food insecure};
  \node[anchor=west, color=AccentGold] at (104,7.4) {Child wasting};
  \node[anchor=west, color=HeadingBlue] at (104,6.95) {Safe water};
\end{tikzpicture}
\end{center}

\subsection{Questions}

Answer all parts. For Question 4, answer either \term{4(a)} or \term{4(b)}.

\begin{enumerate}
  \item \exam{State} two contrasts in food security or health vulnerability
  shown in Source B. \markpts{2}

  \item \exam{Describe} two patterns shown by the food insecurity, safe water,
  and health-risk data in Source B. \markpts{4}

  \item \exam{Explain} two ways a climate shock can worsen food insecurity or
  health. \markpts{4}

  \item Answer either 4(a) or 4(b). \markpts{10}
  \begin{enumerate}[label=\textbf{\alph*.},leftmargin=1.6em]
    \item Using one or more named examples, \exam{examine} the links between
    malnutrition and disease.

    \item Using one or more named examples, \exam{evaluate} the success of
    strategies to improve food security after environmental or economic shocks.
  \end{enumerate}
\end{enumerate}

\end{document}
