\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.68in]{geometry}
\usepackage{array}
\usepackage{tabularx}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[hidelinks]{hyperref}

\hypersetup{
  pdftitle={Geography Standard Level Paper 1 Practice Simulation Set 2 Answer Booklet - 2026-05-07},
  pdfauthor={IB Geography SL Preparation}
}

\setlength{\parskip}{0.38em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.16}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\normalsize\bfseries}
  {}{0pt}{}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[C]{-- \thepage\ --}
\fancyhead[R]{Answer booklet}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\newcommand{\blankbox}[1]{%
\noindent\fbox{%
  \begin{minipage}[t][#1][t]{0.965\textwidth}
  \vspace{0.18em}\rule{\linewidth}{0pt}
  \end{minipage}
}}

\begin{document}

\thispagestyle{empty}
\begin{center}
{\Large Geography}\\[0.35em]
Standard level\\[0.35em]
Paper 1\\[0.8em]
Practice simulation set 2\\[1.6em]
\textbf{Answer booklet}
\end{center}

\vspace{1.5em}

\begin{tabularx}{\textwidth}{L{0.28\textwidth}Y}
Candidate name & \rule{0.66\linewidth}{0.4pt} \\[1.2em]
Session number & \rule{0.66\linewidth}{0.4pt} \\[1.2em]
Date & \rule{0.66\linewidth}{0.4pt} \\
\end{tabularx}

\vspace{1.5em}

\textbf{Instructions}
\begin{itemize}
  \item Write your answers in the spaces provided.
  \item For each option, answer the structured question and one extended-response question.
  \item The maximum mark for this examination paper is \textbf{[40 marks]}.
\end{itemize}

\vfill
\noindent 5 pages \hfill Practice simulation

\newpage

\section{Option F -- Food and health}

\subsection{11.}

\textbf{(a) (i)} Identify the region with the highest percentage of
food-insecure households. \hfill \textbf{[1]}

\blankbox{1.35cm}

\textbf{(ii)} State the percentage of households reporting child wasting in
the dryland smallholder region. \hfill \textbf{[1]}

\blankbox{1.35cm}

\textbf{(b)} Outline one way in which an extended dry season can reduce safe
water access in rural communities. \hfill \textbf{[2]}

\blankbox{2.45cm}

\textbf{(c)} Explain how a climate shock can lead to changes in:

\textbf{(i)} child wasting in vulnerable rural households; \hfill \textbf{[3]}

\blankbox{4.4cm}

\textbf{(ii)} food prices in urban low-income districts. \hfill \textbf{[3]}

\blankbox{4.4cm}

\newpage

\subsection{12.}

\textbf{Answer either part (a) or part (b).}

\textbf{Chosen part:} \rule{0.16\textwidth}{0.4pt} \hfill \textbf{[10]}

\vspace{0.2em}

\blankbox{20.8cm}

\newpage

\section{Option G -- Urban environments}

\subsection{13.}

\textbf{(a) (i)} State which zone has the highest share of trips made by
walking or cycling. \hfill \textbf{[1]}

\blankbox{1.35cm}

\textbf{(ii)} Identify which zone has the highest share of trips made by
private car. \hfill \textbf{[1]}

\blankbox{1.35cm}

\textbf{(b)} Outline one reason why high private-car dependence in outer urban
zones may create environmental stress. \hfill \textbf{[2]}

\blankbox{2.45cm}

\textbf{(c)} Explain how the pattern of urban transport use can be affected by:

\textbf{(i)} one economic factor; \hfill \textbf{[3]}

\blankbox{4.4cm}

\textbf{(ii)} one technological factor. \hfill \textbf{[3]}

\blankbox{4.4cm}

\newpage

\subsection{14.}

\textbf{Answer either part (a) or part (b).}

\textbf{Chosen part:} \rule{0.16\textwidth}{0.4pt} \hfill \textbf{[10]}

\vspace{0.2em}

\blankbox{20.8cm}

\end{document}
