\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.7in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftTint}{HTML}{F7F4EE}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftGreen}{HTML}{EEF7F0}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Workbook - 2026-04-25},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.45em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.25}
\setlist[itemize]{leftmargin=1.3em,itemsep=2pt,topsep=3pt}
\setlist[enumerate]{leftmargin=1.45em,itemsep=3pt,topsep=4pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\Large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.1em}{\color{AccentGold}\titlerule[0.8pt]}]

\titleformat{\subsection}
  {\normalfont\large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}April 25 Workbook}
\fancyhead[R]{\sffamily\color{HeadingBlue}IB Geography SL}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{AccentTeal}{\textbf{#1}}}
\newcommand{\todobox}{\fbox{\rule{0pt}{0.8em}\rule{0.8em}{0pt}}}
\newcommand{\blankbox}[2]{%
\noindent\fcolorbox{RuleGray}{#1}{%
  \begin{minipage}[t][#2][t]{0.965\textwidth}
  \vspace{0.2em}
  \end{minipage}
}}
\newcommand{\casecard}[2]{%
\fcolorbox{RuleGray}{#1}{%
  \begin{minipage}[t][5.15cm][t]{0.455\textwidth}
  \sffamily\bfseries\textcolor{HeadingBlue}{#2}\par
  \vspace{0.25em}
  Issue and causes:\par\vspace{0.38cm}
  Precise fact or statistic:\par\vspace{0.38cm}
  Impact:\par\vspace{0.38cm}
  Stakeholder or response:\par\vspace{0.38cm}
  Evaluation limitation:\par
  \end{minipage}
}}

\begin{document}

\begin{center}
{\sffamily\bfseries\color{HeadingBlue}\LARGE April 25 Urban Environments Workbook}\\[0.25em]
{\sffamily\color{AccentTeal}\large Recall, Timed Practice, Error Log, And Precision Drill}
\end{center}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Use with the April 25 study pack.}}
Print this workbook or write directly into the PDF after conversion. Complete
the pages in order: \term{30 min recall}, \term{30 min practice},
\term{20 min marking}, then \term{10 min drill}.
\end{minipage}
}

\section{Session Checklist}

\begin{tabularx}{\textwidth}{L{0.12\textwidth}Y L{0.18\textwidth}}
\toprule
\textbf{Done} & \textbf{Task} & \textbf{Time} \\
\midrule
\todobox & Closed-book Urban environments recall & 30 min \\
\todobox & Timed Paper 1 structured practice & 30 min \\
\todobox & Marking and error log & 20 min \\
\todobox & Numbers, place names, and definitions drill & 10 min \\
\bottomrule
\end{tabularx}

\section{30-Minute Recall}

\subsection{Four-Cluster Map From Memory}

Fill every box before checking the study pack.

\begin{tabularx}{\textwidth}{Y Y}
\toprule
\textbf{\textcolor{HeadingBlue}{Cluster 1}} &
\textbf{\textcolor{HeadingBlue}{Cluster 2}} \\
\midrule
\rule{0pt}{4.3cm} & \rule{0pt}{4.3cm} \\
\bottomrule
\end{tabularx}

\vspace{0.65em}

\begin{tabularx}{\textwidth}{Y Y}
\toprule
\textbf{\textcolor{HeadingBlue}{Cluster 3}} &
\textbf{\textcolor{HeadingBlue}{Cluster 4}} \\
\midrule
\rule{0pt}{4.3cm} & \rule{0pt}{4.3cm} \\
\bottomrule
\end{tabularx}

\subsection{Urban Land-Use Pattern Sketch}

Sketch a simple city pattern. Label at least five of these: CBD, inner city,
suburb, rural-urban fringe, transport corridor, industrial zone, informal
settlement, green space, edge city.

\blankbox{white}{7.2cm}

\subsection{Pattern To Process Chains}

Complete three chains. Each chain should move from observation to explanation.

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.21\textwidth}Y Y Y}
\toprule
\textbf{Pattern} & \textbf{Likely process} & \textbf{Impact} & \textbf{Management or evaluation} \\
\midrule
\rule{0pt}{1.4cm} & & & \\
\rule{0pt}{1.4cm} & & & \\
\rule{0pt}{1.4cm} & & & \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\newpage

\subsection{Urbanization And Migration Recall}

\begin{tabularx}{\textwidth}{Y Y}
\toprule
\textbf{\textcolor{HeadingBlue}{Urbanization definition and causes}} &
\textbf{\textcolor{HeadingBlue}{Migration push and pull factors}} \\
\midrule
\rule{0pt}{5.3cm} & \rule{0pt}{5.3cm} \\
\bottomrule
\end{tabularx}

\vspace{0.7em}

\begin{tabularx}{\textwidth}{Y Y}
\toprule
\textbf{\textcolor{HeadingBlue}{Pressure created by urban growth}} &
\textbf{\textcolor{HeadingBlue}{Opportunity created by urban growth}} \\
\midrule
\rule{0pt}{5.3cm} & \rule{0pt}{5.3cm} \\
\bottomrule
\end{tabularx}

\subsection{Named Example Fact Cards}

Fill at least three cards. Use class notes for exact statistics.

\casecard{SoftTint}{Detroit, USA}
\hfill
\casecard{SoftBlue}{Mumbai or Dharavi}

\vspace{0.55em}

\casecard{SoftBlue}{Curitiba or sustainable-city example}
\hfill
\casecard{SoftTint}{Your class example}

\newpage

\section{30-Minute Timed Practice}

\subsection{Source A}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.19\textwidth}Y Y Y Y}
\toprule
\textbf{Zone} & \textbf{Dominant land use} & \textbf{Population density} &
\textbf{Green space} & \textbf{Key issue} \\
\midrule
Central business district & Offices, retail, transport terminals, apartments
above shops & Medium at night, very high by day & Low & High land value and
traffic congestion \\
Inner city & Older housing, former industry, new apartments, cultural venues &
High & Low to medium & Reinvestment and possible displacement \\
Suburbs & Residential areas, schools, local services, retail parks & Medium to
low & Medium to high & Long commutes and land consumption \\
Peripheral informal area & Self-built housing, small workshops, informal retail
& Very high & Very low & Weak service access and insecure housing \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsection{Question 1: State Two Features}

\textbf{State two features of the urban land-use pattern shown in Source A.}
\hfill \textbf{2 points}

\blankbox{white}{3.1cm}

\subsection{Question 2: Describe Two Contrasts}

\textbf{Describe two contrasts between the central business district and the
suburbs.} \hfill \textbf{4 points}

\blankbox{white}{5.2cm}

\newpage

\subsection{Question 3: Explain Two Reasons}

\textbf{Explain two reasons why migration can increase pressure on urban
housing or services.} \hfill \textbf{4 points}

\blankbox{white}{7.4cm}

\subsection{Question 4: Named Example}

\textbf{Using Detroit or the named urban example you have studied, examine how
gentrification or regeneration can create both benefits and costs.}
\hfill \textbf{10 points}

\blankbox{white}{11.8cm}

\newpage

\section{20-Minute Mark And Error Log}

\subsection{Score Sheet}

\begin{tabularx}{\textwidth}{L{0.16\textwidth}L{0.18\textwidth}Y}
\toprule
\textbf{Question} & \textbf{Score} & \textbf{One reason for the score} \\
\midrule
1 & \rule{0.08\textwidth}{0.4pt} / 2 & \rule{0.58\textwidth}{0.4pt} \\
2 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
3 & \rule{0.08\textwidth}{0.4pt} / 4 & \rule{0.58\textwidth}{0.4pt} \\
4 & \rule{0.08\textwidth}{0.4pt} / 10 & \rule{0.58\textwidth}{0.4pt} \\
Total & \rule{0.08\textwidth}{0.4pt} / 20 & \rule{0.58\textwidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\subsection{Error Log}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.18\textwidth}Y Y Y}
\toprule
\textbf{Error type} & \textbf{What I wrote or missed} & \textbf{Correct version} & \textbf{Next action} \\
\midrule
\rule{0pt}{1.6cm} & & & \\
\rule{0pt}{1.6cm} & & & \\
\rule{0pt}{1.6cm} & & & \\
\rule{0pt}{1.6cm} & & & \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsection{Rewrite One Weak Answer}

Choose the weakest answer and rewrite it below using the correct command term.

\blankbox{SoftTint}{7.1cm}

\newpage

\section{10-Minute Precision Drill}

\subsection{Definitions}

Write a clean definition in one line. Add a quick example if time allows.

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.25\textwidth}Y}
\toprule
\textbf{Term} & \textbf{Definition from memory} \\
\midrule
Urbanization & \rule{0.86\linewidth}{0.4pt} \\
Migration & \rule{0.86\linewidth}{0.4pt} \\
Suburbanization & \rule{0.86\linewidth}{0.4pt} \\
Counter-urbanization & \rule{0.86\linewidth}{0.4pt} \\
Reurbanization & \rule{0.86\linewidth}{0.4pt} \\
Megacity & \rule{0.86\linewidth}{0.4pt} \\
Gentrification & \rule{0.86\linewidth}{0.4pt} \\
Deindustrialization & \rule{0.86\linewidth}{0.4pt} \\
Urban sprawl & \rule{0.86\linewidth}{0.4pt} \\
Informal settlement & \rule{0.86\linewidth}{0.4pt} \\
Urban environmental stress & \rule{0.86\linewidth}{0.4pt} \\
Urban social stress & \rule{0.86\linewidth}{0.4pt} \\
Sustainable urban management & \rule{0.86\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\subsection{Numbers And Places}

\begin{tabularx}{\textwidth}{L{0.22\textwidth}Y Y}
\toprule
\textbf{Place} & \textbf{Number or exact detail} & \textbf{How I would use it in an answer} \\
\midrule
Detroit & \rule{0.86\linewidth}{0.4pt} & \rule{0.86\linewidth}{0.4pt} \\
Detroit district or project & \rule{0.86\linewidth}{0.4pt} & \rule{0.86\linewidth}{0.4pt} \\
Informal settlement example & \rule{0.86\linewidth}{0.4pt} & \rule{0.86\linewidth}{0.4pt} \\
Sustainable-city example & \rule{0.86\linewidth}{0.4pt} & \rule{0.86\linewidth}{0.4pt} \\
Your class example & \rule{0.86\linewidth}{0.4pt} & \rule{0.86\linewidth}{0.4pt} \\
\bottomrule
\end{tabularx}

\section{Mastery Update}

Rate yourself after marking. Be strict enough that the score is useful.

\begin{tabularx}{\textwidth}{Y L{0.15\textwidth}Y}
\toprule
\textbf{Skill} & \textbf{Rating} & \textbf{Evidence} \\
\midrule
I can recall the Urban environments four-cluster map. & \_\_ /5 & \\
I can explain land-use patterns using process language. & \_\_ /5 & \\
I can connect migration to housing and service pressure. & \_\_ /5 & \\
I can use Detroit or another named example with precise evidence. & \_\_ /5 & \\
I can answer a structured question under time. & \_\_ /5 & \\
\bottomrule
\end{tabularx}

\vspace{0.7em}

\noindent\colorbox{SoftGreen}{%
\begin{minipage}{0.965\textwidth}
\sffamily\raggedright
\textbf{\textcolor{HeadingBlue}{Carry forward to Monday 2026-04-27.}}
My weakest Urban environments item is:
\rule{0.54\textwidth}{0.4pt}
\end{minipage}
}

\end{document}
