\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.68in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage[normalem]{ulem}

\definecolor{TextInk}{HTML}{233142}
\definecolor{HeadingBlue}{HTML}{1D3557}
\definecolor{AccentTeal}{HTML}{2A7F92}
\definecolor{AccentGold}{HTML}{C58B4E}
\definecolor{RuleGray}{HTML}{D6E0E5}
\definecolor{TableStripe}{HTML}{F3F8FA}
\definecolor{SoftTint}{HTML}{F7F4EE}
\definecolor{SoftBlue}{HTML}{EEF6F8}
\definecolor{SoftGreen}{HTML}{EEF7F0}
\definecolor{SoftRed}{HTML}{FBF0EF}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=AccentTeal,
  pdftitle={IB Geography SL Case Study Rapid Recall Workbook - 2026-05-01},
  pdfauthor={IB Geography SL Preparation}
}

\color{TextInk}
\setlength{\parskip}{0.42em}
\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.23}
\setlist[itemize]{leftmargin=1.25em,itemsep=2pt,topsep=3pt}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\titleformat{\section}
  {\normalfont\Large\sffamily\bfseries\color{HeadingBlue}}
  {}{0pt}{}
  [\vspace{0.08em}{\color{AccentGold}\titlerule[0.8pt]}]

\titleformat{\subsection}
  {\normalfont\large\sffamily\bfseries\color{AccentTeal}}
  {}{0pt}{}

\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\fancyhead[L]{\sffamily\color{AccentTeal}May 1 Workbook}
\fancyhead[R]{\sffamily\color{HeadingBlue}IB Geography SL}
\fancyfoot[C]{\sffamily\color{HeadingBlue}\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\term}[1]{\textcolor{HeadingBlue}{\textbf{#1}}}
\newcommand{\exam}[1]{\textcolor{AccentTeal}{\textbf{#1}}}
\newcommand{\todobox}{\fbox{\rule{0pt}{0.8em}\rule{0.8em}{0pt}}}
\newcommand{\smallline}{\rule{0.96\linewidth}{0.4pt}}

\begin{document}

\begin{center}
{\sffamily\bfseries\color{HeadingBlue}\LARGE May 1 Case-Study Rapid Recall Workbook}\\[0.25em]
{\sffamily\color{AccentTeal}\large Place $\rightarrow$ Issue $\rightarrow$ Causes $\rightarrow$ Impacts $\rightarrow$ Response $\rightarrow$ Evaluation}
\end{center}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Use with the May 1 study pack.}}
The plan requires \term{25 minutes} to create a rapid-recall sheet. Work
closed-book first. Add only statistics you can remember accurately; mark
uncertain facts with a question mark and check them after the timed block.
\end{minipage}
}

\section{Session Checklist}

\begin{tabularx}{\textwidth}{L{0.11\textwidth}Y L{0.17\textwidth}}
\toprule
\textbf{Done} & \textbf{Task} & \textbf{Time} \\
\midrule
\todobox & Closed-book dump of case names and statistics & 4 min \\
\todobox & Select the best Urban and Resources examples & 4 min \\
\todobox & Fill four six-link case-study rows & 9 min \\
\todobox & Add one statistic and one limitation per case & 4 min \\
\todobox & Cover the notes and retrieve each chain aloud & 3 min \\
\todobox & Flag uncertain facts for checking & 1 min \\
\bottomrule
\end{tabularx}

\section{Minute 0--4: Closed-Book Dump}

Write from memory before opening the study pack.

\begin{tabularx}{\textwidth}{L{0.23\textwidth}Y}
\toprule
\textbf{Prompt} & \textbf{My recall} \\
\midrule
Urban case names & \smallline \\[0.75em]
Urban statistics & \smallline \\[0.75em]
Urban responses or management & \smallline \\[0.75em]
Resources case names & \smallline \\[0.75em]
Resources statistics & \smallline \\[0.75em]
Resources stewardship responses & \smallline \\[0.75em]
\bottomrule
\end{tabularx}

\section{Minute 4--17: Six-Link Case Rows}

\subsection{Urban Environments}

\begin{tabularx}{\textwidth}{L{0.15\textwidth}Y Y}
\toprule
\textbf{Link} & \textbf{Case 1: Detroit} & \textbf{Case 2: backup / class case} \\
\midrule
Place & & \\
Issue & & \\
Causes & & \\
Impacts & & \\
Response & & \\
\uline{Evaluation} & & \\
\bottomrule
\end{tabularx}

\vspace{0.55em}

\subsection{Resources}

\begin{tabularx}{\textwidth}{L{0.15\textwidth}Y Y}
\toprule
\textbf{Link} & \textbf{Case 3: China / e-waste} & \textbf{Case 4: Singapore / water} \\
\midrule
Place & & \\
Issue & & \\
Causes & & \\
Impacts & & \\
Response & & \\
\uline{Evaluation} & & \\
\bottomrule
\end{tabularx}

\section{Post-Timer Carry-Forward Rows}

Use these after the 25-minute timed block, or during review if either case-bank
slot is still empty. They address the April 30 carry-forward gaps without
changing today's Urban + Resources timing.

\begin{tabularx}{\textwidth}{L{0.15\textwidth}Y Y}
\toprule
\textbf{Link} & \textbf{Food and health: Haiti cholera} & \textbf{Climate: Bangladesh cyclones} \\
\midrule
Place & & \\
Issue & & \\
Causes & & \\
Impacts & & \\
Response & & \\
\uline{Evaluation} & & \\
\bottomrule
\end{tabularx}

\section{Minute 17--21: Statistic + Limitation Check}

\rowcolors{2}{TableStripe}{white}
\begin{tabularx}{\textwidth}{L{0.21\textwidth}Y Y}
\toprule
\textbf{Case} & \textbf{One statistic I can use accurately} & \textbf{One limitation / evaluation point} \\
\midrule
Detroit & & \\
Curitiba or other urban case & & \\
China / e-waste & & \\
Singapore / water & & \\
Haiti cholera & & \\
Bangladesh cyclones & & \\
\bottomrule
\end{tabularx}
\rowcolors{2}{}{}

\section{Minute 21--24: Cover-And-Recall Test}

Cover the case rows. Write only the chain keywords from memory.

\begin{tabularx}{\textwidth}{L{0.2\textwidth}Y}
\toprule
\textbf{Case} & \textbf{Memory chain} \\
\midrule
Detroit & \smallline \\[0.8em]
Curitiba / urban backup & \smallline \\[0.8em]
China / e-waste & \smallline \\[0.8em]
Singapore / water & \smallline \\[0.8em]
\bottomrule
\end{tabularx}

\section{Minute 24--25: Accuracy Flag}

\begin{tabularx}{\textwidth}{L{0.24\textwidth}Y Y}
\toprule
\textbf{Uncertain detail} & \textbf{Why I am unsure} & \textbf{Where to check} \\
\midrule
 & & \\
 & & \\
 & & \\
\bottomrule
\end{tabularx}

\newpage

\section{Prompt Practice}

Use these only after completing the rapid-recall sheet.

\subsection{Urban Prompts}

\noindent\colorbox{SoftTint}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt A.}}
\exam{Evaluate} the success of urban regeneration in reducing social stress in
one named urban area.
\end{minipage}
}

\vspace{0.45em}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt B.}}
\exam{Explain} how one strategy can make an urban area more sustainable.
\end{minipage}
}

\subsection{Resource Prompts}

\noindent\colorbox{SoftTint}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt C.}}
\exam{Analyse} how resource consumption can create environmental problems in
one or more places.
\end{minipage}
}

\vspace{0.45em}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt D.}}
\exam{Evaluate} the success of one stewardship strategy in improving resource
security.
\end{minipage}
}

\vspace{0.45em}

\noindent\colorbox{SoftTint}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt E.}}
\exam{Explain} how water and sanitation can influence the spread of disease in
one named place.
\end{minipage}
}

\vspace{0.45em}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Prompt F.}}
\exam{Evaluate} the success of one strategy used to reduce climate
vulnerability.
\end{minipage}
}

\section{One-Paragraph Planning Space}

\begin{tabularx}{\textwidth}{L{0.2\textwidth}Y}
\toprule
\textbf{Need} & \textbf{Notes} \\
\midrule
Prompt chosen & A \quad B \quad C \quad D \quad E \quad F \\
Command term & \smallline \\[0.8em]
Case selected & \smallline \\[0.8em]
Statistic & \smallline \\[0.8em]
Explanation chain & \smallline \\[0.8em]
\uline{Evaluation} & \smallline \\[0.8em]
\bottomrule
\end{tabularx}

\section{Fast Answer Frames}

\noindent\colorbox{SoftGreen}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Urban.}}
In \rule{0.18\textwidth}{0.4pt}, \rule{0.20\textwidth}{0.4pt} developed
because \rule{0.25\textwidth}{0.4pt}. This caused
\rule{0.25\textwidth}{0.4pt}. The response was
\rule{0.22\textwidth}{0.4pt}; however, its success was limited by
\rule{0.25\textwidth}{0.4pt}.
\end{minipage}
}

\vspace{0.55em}

\noindent\colorbox{SoftBlue}{%
\begin{minipage}{0.965\textwidth}
\sffamily
\textbf{\textcolor{HeadingBlue}{Resources.}}
In \rule{0.18\textwidth}{0.4pt}, resource insecurity was caused by
\rule{0.25\textwidth}{0.4pt}. The impact was
\rule{0.25\textwidth}{0.4pt}. Stewardship through
\rule{0.22\textwidth}{0.4pt} improved security, but it was limited by
\rule{0.25\textwidth}{0.4pt}.
\end{minipage}
}

\section{Error Log}

\begin{tabularx}{\textwidth}{L{0.22\textwidth}Y Y}
\toprule
\textbf{Problem} & \textbf{What happened} & \textbf{Fix before May 2 Paper 1 mock} \\
\midrule
Forgot statistic & & \\
Weak cause chain & & \\
Response too vague & & \\
No evaluation & & \\
Used a case that did not fit prompt & & \\
\bottomrule
\end{tabularx}

\end{document}
